---
title: "Project_1"
author: "Ana OR"
date: "8/4/2020"
output: html_document
---

## Topic Background

Each of our team members originally selected a few datasets that we compared as a group.  Our group decided to explore a dataset Ana found related to the topic of racial and ethnic diversity at two-year and four-year colleges and universities in the United States.  We intially thought to compare diversity at two-year community colleges versus 4-year schools and also were curious about the variations in diversity between the two that could be found comparatively across the country.

As we were exploring the dataset we identified a number of named colleges and universities that are known as "for-profit" institutions. Some of these institutions have been the subject of recent lawsuits regarding predatory (Halperin, 2020; Legal Services Center, 2020; Redman, 2020).  The high number of current lawsuits regarding for-profit colleges' predatory behaviors are a result of the actions of current Education Secretary Betsy DeVos.  In 2019, DeVos "repealed an Obama-era regulation that sought to crack down on for-profit colleges and universities that produced graduates with no meaningful job prospects and mountains of student debt they could not hope to repay" (Green, 2019. para. 1).  Another action that DeVos took was to deny debt forgiveness to students who had been prey to for-profit predatory educational institutions (Lobosco, 2019; Turner, 2019).  We were aware of past and recent news articles describing the marketing strategies of these colleges that were aimed at students of color, low-income students, immigrant communities and students who are first in their family to go to college (Bonadies et al. 2018; Conti, 2019; Voorhees, 2019).  We thought it would be interesting to examine our dataset for consistencies or inconsistencies with these reports.

## Initial Questions

1. How do racial and ethnic demographics vary between not-for-profit institutions and for-profit colleges?

2. Do the For-Profit Institutions have a higher proportion of students of color than other community colleges and four-year colleges, thus providing support to the argument that For-Profit colleges target students of color for enrollment?

## Definitions:
For-Profit Institution:

Not-for Profit Institution:

## Webscraping and Randomly Selecting Sample of 35 For-Profit Colleges

Our dataset did not include information on whether colleges were for-profit.

In order to determine which of the schools on our list were for-profit institutions, we decided to find a website with a list of for-profit institutions, scrape the data from that website, to create a second dataset with a list of for profit schools.  

# Following Web-Scrapping turtorial to scrape Wikipedia website

# Loading Libraries
```{r}
#Loading the rvest package
library('rvest')
library(tidyverse)
library(dplyr)
```

```{r}
#Specifying the url for desired website to be scraped
url <- 'https://en.wikipedia.org/wiki/List_of_for-profit_universities_and_colleges'

#Reading the HTML code from the website
webpage <- read_html(url)

#Using CSS selectors to scrape the rankings section
for_profit_html <- html_nodes(webpage,'li a')
#Converting the ranking data to text
for_profit <- html_text(for_profit_html)

#Let's have a look at the rankings
head(for_profit)
```

# Viewing new index
```{r}
# for_profit
```

```{r}
for_profit_new <- for_profit[25:215]
head(for_profit_new)
```
## cleaning our for-profit list
```{r}
for_profit_new[167:168] <- NA
for_profit_new
```

```{r}
final_for_profit <- na.omit(for_profit_new)
```

```{r}
#final_for_profit
 df <-data_frame(final_for_profit)
df
```

The webscraping exercise produced a list of over 1,000 schools.  We decided to take a random sample of n=35  schools on that list, which also appeared in our dataset. We selected 35 schools so that we would comply with the Central Limit Theorem.  The result of the randomization we ran on the webscraped data is below.  Each of us compared a portion of the list with our dataset to ensure that all of our selections were present on both lists.  We ran multiple randomizations until we had a consistent list with 35 schools appearing on both datasets.

```{r}
sample_n(df,35)
```

# Jennifer
Spencerian College
- Spencerian College at Louisville (Ky.) 
- Spencerian College at Lexington (Ky.
Aspen University	
American Public University	
- American Public University system
Western International University	
NewSchool of Architecture and Design	
Schiller International University	
National Paralegal College	
West Coast University	
- West Coast University at Ontario
- West Coast University trasound Institute 
- West Coast University at Dallas
- West Coast University at Miami
- West Coast University -Los Angeles
- West Coast University -Orange County
Mildred Elley	
- Mildred Elley
- Mildred Elley at New York City
Brookline College	
- Brookline College at Tucson
- Brookline College at Tempe (Ariz.)
- Brookline College at Albuquerque
- Brookline College at Phoenix
Grand Canyon University	

#Ana
Walden University	
Blue Cliff College	
Neumont University	
University of Pheonix
Stevens-Henager College	
DeVry University	
Pioneer Pacific College	
Capella University	
Grantham University	   
Stratford University	
Redstone College	
National College	

# Tiffany
Strayer University	
Lincoln Tech	
Full Sail University	
Rocky Mountain College of Art and Design	
Minneapolis Business College	
Fashion Institute of Design & Merchandising	
Paier College of Art
Vista College	
Centura College	
Rasmussen College
Redstone College	
Fortis College	
Bay State College	



## Data Wrangling & Analysis

The dataset was retrieved from Kaggle, available here: https://www.kaggle.com/jessemostipak/college-tuition-diversity-and-pay?select=diversity_school.csv  

The dataset is composed of five separate .csv files addressing school diversity, historical tuition rates, salary protential, the cost of tuition in 2016 and the income of students compared to tuition.  We mostly relied upon the school diversity file. 


### Setting working directory
```{r}
setwd("/Users/tiffanyking/Documents/R Studio Project 1 ") 
diversity <- read_csv("diversity2.csv")
```
### Viewing data
```{r}
head(diversity)
```

## Cleaning the Data

There were numerous "tidying" exercises that we undertook to clean our data and make the webscraping dataset and the diversity dataset comparable. We also created a new column calculating the percent of students in attendance based on race and ethnicity and a new binary column identifying the selected for-profit and not-for-profit schools represented by "1" and "0" respectively.

### Replacing White spaces with underscores
```{r}
#gsub(" ", "_", diversity$name)
#gsub(" ", "_", diversity$category) 
```

### Creating new column for percent enrollment
```{r}
diversity_new <- diversity %>%  mutate(enrollment_percentage = enrollment/ total_enrollment*100)
```

```{r}
head(diversity_new)
```
```{r}
diversity_new %>% filter(name=="Strayer University")
```


```{r}
unique(diversity_new$name)
```


```{r}
devry <- filter(diversity_new, str_detect(name, "DeVry"))
pheniox <- filter(diversity_new, str_detect(name,"Pheniox"))
strayer <- filter(diversity_new, str_detect(name, "Strayer"))
```

```{r}
str(diversity_new)
```
## Option to Change names for multiple locations into one name

```{r}
diversity_new$name[diversity_new$name == "Spencerian College at Louisville (Ky.)"] <- "Spencerian_College"
```

```{r}
diversity_new$name[diversity_new$name == "Spencerian College at Lexington (Ky.)"] <- "Spencerian_College"
```

```{r}
diversity_new$name %>% 
  if(name %in% "Mildred Elley") {
  name == "Mildred_Elley"
  }  
```



### 35 Randomly selected not-for-profit educational institutions

For consistency of methodology we decided to select 35 schools from the not-for-profit educational institutions to compare to the 35 randomly selected for-profit schools.  

```{r}
sample_n(diversity_new,35)
```

# Jennifer
University of Idaho
Southwestern Community College (Iowa)
Clover Park Technical College
Clark College
City University of New York Hunter College
University of Houston
- University of Houston
- University of Houston-Downtown
- University of Houston-Clear Lake
- University of Houston-Victoria
University of Colorado 
- University of Colorado at Colorado Springs
- University of Colorado at Boulder
- University of Colorado at Denver
University of Massachusetts
- University of Massachusetts at Dartmouth
- University of Massachusetts at Worcester
- University of Massachusetts at Amherst
- University of Massachusetts at Lowell
- University of Massachusetts at Boston
Arizona College
- Arizona College at Glendale
- Arizona College at Mesa
San Diego City College

#Ana
State University of New York College at Purchase
Eastern Shore Community College
Southern Oregon University
Santa Monica College
Adirondack Community College
East Georgia State College
Smith College
East Tennessee State University
Austin Community College
Tompkins Cortland Community College

#Tiffany
University of Hawaii Hawaii Community College
Burlington College
Pennsylvania State University - Harrisburg
University of Minnesota
University of Central Oklahoma
Southern Virginia University
Pennsylvania Highlands Community College
Arizona State University
ITT Technical Institute at Knoxville
Colgate University

University of Pittsburg

## Create a new column to identify the randomly selected for-profit colleges sample.

```{r}
#diversity_new2 <- diversity %>%  mutate(for_profit = 
#if name == "Spencerian_College" )
```






```{r, include=FALSE}
### Produces all False Columns
diversity_new2 <- diversity_new %>% mutate(forProfit = str_detect(name =='Spencerian_College', "1"))
```

```{r, include=FALSE}
diversity_new2$forProfit[47834]
```
```{r, include=FALSE}
sum(diversity_new2$forProfit == FALSE)
```


### Produces 1 for For-Profit Random Selections and 0 for the rest

```{r}
diversity_new3 <- diversity_new %>% mutate(diversity_new,forProfit = ifelse(name=="Spencerian_College", 1,
     ifelse(name=="Aspen University", 1, 
        ifelse(name=="American Public University system", 1, 0))))
```


```{r}
diversity_new3$forProfit[47834] # Check on Spencerian_College entry
diversity_new3$forProfit[24533] # Check on Aspen University
diversity_new3$forProfit[123] # Check on American Public University system
```

##**Tiffany's FOR PROFIT Schools**
```{r}
diversity_new$name[grepl('Strayer University',diversity_new$name)]<- 'Strayer_Univ.'
```

```{r}
diversity_new$name[grepl('Lincoln Tech',diversity_new$name)]<- 'Lincoln_Tech' 
```

```{r}
diversity_new$name[grepl('Fashion Institute of Design & Merchandising',diversity_new$name)]<- 'Fashion_Institute' 
```

```{r}
diversity_new$name[grepl('Centura College' ,diversity_new$name)]<- 'Centura_College'
```

```{r}
diversity_new$name[grepl('Rasmussen College' ,diversity_new$name)]<- 'Rasmussen_College'
```

```{r}
diversity_new$name[grepl('Fortis College' ,diversity_new$name)]<- 'Fortis_College'
```


##**Tiffany's Non Profit Schools**

```{r}
diversity_new$name[grepl('Pennsylvania State University',diversity_new$name)]<- 'Penn_State_Univ.'
```

```{r}
diversity_new$name[grepl('University of Minnesota',diversity_new$name)]<- 'Univ._of_Minn'
```

```{r}
diversity_new$name[grepl('Arizona State University',diversity_new$name)]<- 'Arizona_State' 
```

```{r}
diversity_new$name[grepl('ITT Technical Institute',diversity_new$name)]<- 'ITT_Tech'
```


## **References**

Bonadies, G.G., Rovenger, J., Connor, E., Shum, B. & Merrill, T. (2018, Jul. 30). For-Profit Schools' Predatory Practices and Students of Color:  A Mission to Enroll Rather than Educate, *Harvard Law Review Blog*. https://blog.harvardlawreview.org/for-profit-schools-predatory-practices-and-students-of-color-a-mission-to-enroll-rather-than-educate/  

Conti, A. (2019, Sep. 10).  How For-Profit Colleges Have Targeted and Taken Advantage of Black Students. *Vice*. https://www.vice.com/en_us/article/bjwj3d/how-for-profit-colleges-have-targeted-and-taken-advantage-of-black-students

Green, E.L. (2019, Jun. 28).  DeVos Repeals Obama-Era Rule Cracking down on For-Profit Colleges, *New York Times*. https://www.nytimes.com/2019/06/28/us/politics/betsy-devos-for-profit-colleges.html

Halperin, D. 22 States Sue DeVos to Overturn Anti-Student Rule.  *Republic Report*. https://www.republicreport.org/2020/22-states-sue-devos-to-overturn-anti-student-rule/

Legal Services Center(2020), *Project on Predatory Student Lending: Cases*, Harvard Law School.  https://predatorystudentlending.org/cases/

Lobosco, K. (2019, Jul. 23).  For-profit college students are waiting 958 days for loan relief, *CNN*.  https://www.cnn.com/2019/07/23/politics/betsy-devos-loan-forgiveness-for-profit-college-students/index.html

Redman, H. (2020, Jun. 27).  AG Sues Department of Education Over For-Profit College Rules.  *Urban Milwaukee*.  https://urbanmilwaukee.com/2020/06/27/ag-sues-department-of-education-over-for-profit-college-rules/

Turner, C. (2019, Nov. 14).  Devos Refuses to Forgive student Debt For Those DeFrauded by For-Profit Colleges, *All Things Considered, NPR*. https://www.npr.org/2019/11/14/779465130/devos-refuses-to-forgive-student-debt-for-those-defrauded-by-for-profit-colleges

Voorhees, K. (2019, Oct. 17).  Civil Rights Groups: For-Profit Colleges Exploit Black and Latino Students. *The Leadership Conference Education Fund*. https://civilrights.org/edfund/2019/10/17/civil-rights-groups-for-profit-colleges-exploit-black-and-latino-students/